---
title: "Problem Set 1:Getting Started Key"
author: "Claire Duquennois"
output:
  html_document:
    number_sections: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Group Member 1: Chean Shea**

**Group Member 2: Mayowa Idowu**

**Group Member 3: Milan Stefanelli**

**Group Member 4: Kiersten Kochanowski**



# Omitted Variable Bias Simulation (Group excercise)

```{r sim1,  warning=FALSE}
library(MASS)
library(ggplot2)

out <- as.data.frame(mvrnorm(1000, mu = c(0,0), 
                     Sigma = matrix(c(1,0.8,0.8,1), ncol = 2), 
                     empirical = TRUE))
cor(out)
plot(out)

```

Next I generate a randomly distributed error term and I calculate the outcome variable which depends on both V1 and V2 and some noise:

$$Y=\beta_1V_1+\beta_2V_2+\epsilon$$

```{r sim2}
out$error<-rnorm(1000, mean=0, sd=1)

#The data generating process
B1<-3
B2<-6

out$Y<-out$V1*B1+out$V2*B2+out$error
```

**TO DO: For the questions below write the needed code and a written response to the question. **

**Question 1: Write a chunk in which you regress $Y$ on $V_1$ and $V_2$. Are your estimates of $\beta_1$ and $\beta_2$ biased?**

**Answer:** No, our data was randomly generated (within parameters) so their coefficients are not bias. This is also affirmed by the fact that the V1 and V2 coefficients in our regression model are almost identical to the one B1 and B2 generated above.

```{r include=FALSE}
library(stargazer)
```


```{r}
View(out)
M1 <- lm(data=out, Y ~ V1 + V2)

stargazer(M1, type="text", header=FALSE, title="Regression Coefficients", 
                     style="qje", 
                     covariate.labels = c("B1", "B2", "Intercept"),
                     column.labels = c("Model #1"))
```


**Question 2: Write a chunk in which you regress $Y$ on $V_1$ only. Is your estimate of $\beta_1$ biased?**

**Answer:** Yes, our beta 1 (V1) coefficient is biased due to omitted variable bias. By omitting V2, our beta 1 coefficient has upward bias (larger than it would be with V2 in the model: 7.81 vs 2.93). 

```{r}
M2 <- lm(data=out, Y~V1)

library(stargazer)
stargazer(M2, M1, type="text", header=FALSE, title="Comparison of Regression Coefficients (M2 & M1)", 
                     style="qje", 
                     covariate.labels = c("B1", "B2", "Intercept"),
                     column.labels = c("Model #2", "Model #1"))
```

**Question 3: Generate a new variable $Y_{adj}$ such that $Y_{adj}=Y-\beta_2*V_2$. Then regress $Y_{adj}$ on $V_1$. Is your estimate of $\beta_1$ biased? Can you explain why/why not?**

**Answer:** No, beta 1 is not biased in this model. We can see that the coefficient is very similar to model #1 where beta two was also included (3.01 vs 2.93). We can also affirm this mathematically given the relationship below

Y-B2V2 = B1V1 (Model #3), which is equal to Y=B1V1 + B2V2 (Model #1)

```{r}
Y_adj <- ((out$Y)-(B2*out$V2))
M3 <- lm(data=out, Y_adj ~ V1)
stargazer(M3, M1, type="text", header=FALSE, title="Comparison of Regression Coefficients (M3 & M1)", 
                     style="qje", 
                     covariate.labels = c("B1", "B2", "Intercept"),
                     column.labels = c("Model #3", "Model #1"))
```

**Question 4: The code below generates a scatter plot and regression line for the relationship between $V_1$ and $Y$ as well as $V_1$ and $Y_{adj}$. Submit an improved visualization of this data. Hint: you will need to delete the # to get the code to run**

```{r key 4}

plotted<-ggplot(out, aes(V1, y = value, color = variable)) + 
   geom_point(aes(y = Y, col = "Y")) + 
   geom_point(aes(y = Y_adj, col = "Y_adj"))+
   geom_smooth(method='lm', aes(y = Y, col = "Y"))+
   geom_smooth(method='lm', aes(y = Y_adj, col = "Y"))

plotted 
```

\newpage
Loading packages to run plot_coefs
```{r include=FALSE}
install.packages("ggstance", repos = "http://cran.us.r-project.org" )
install.packages("jtools", repos = "http://cran.us.r-project.org")
install.packages("broom.mixed", repos = "http://cran.us.r-project.org")
library(GGally)
library(jtools)
library(broom.mixed)
```

**New Data Visualization**
We interpreted the provided visual to be describing the relationship between V1 and Y with and without omitted variable bias (OVB). A simpler way to put OVB into perspective is by illustrating the difference in the V1 coefficient (and its relationship to Y) when it is and isn't biased by missing variable V2. 
```{r}
data_vis <- plot_coefs(M2, M3, omit.coefs =c("(Intercept)","V2"), model.names=c("OVB", "No OVB"))
data_vis2 <- data_vis + coord_flip()
data_vis2
```




